References for PloS paper

### 2014-02-22 image_analysis_paper/math_description.tex
added math description

\usepackage{graphicx}
\graphicspath{ {paper_figures/} }
\usepackage{subcaption}

\usepackage{forloop}

@@

\section{Introduction}

% only images
We have cross-sectional data. When there is just a few snapshots and differences between them are obvious, such as the overall size or some other morphological feature, ordering is easy. We can point to "fish" as our example.

But there is more and more studies with lots of snapshots, lots of noise, and subtle differences between snapshots. Our example is the dataset from the embryo. We discuss how to align and temporally order cross sectional data in this case.

Maybe figure 1 is fish ordered (B) and unordered (A).

Figure 2 can be the embryo dataset - unordered and unoriented (A) and hand-ordered (by membranes) and hand-oriented (by Dorsal) (B).

This way Figure 2 presents the data and makes it a reference for all future alignments.

@@

% show schematic with images (starfish or zebrafish w/ stripes)

@@

Let $I_1, \dots, I_m \in \mathbb{R}^{n \times n}$ denote the (grayscale) images that we wish to align with respect to translations and rotations.
%
The first step in angular synchronization is to compute the translations and rotations to optimally align each pair of data points. 
%
For each image pair $I_i$ and $I_j$, we would like to solve
\begin{equation}\label{eq:opt_pairwise}
(\theta_{ij}, dx_{ij}, dy_{ij}) = \argmin_{
\begin{matrix}
\theta \in [0, 2\pi) \\ 
dx \in [-\Delta, \Delta]\\ 
dy \in [-\Delta, \Delta]
\end{matrix}
} \|f(I_j, \theta, dx, dy) - I_i \|_F^2.
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm, and $f(I_j, \theta, dx, dy)$ is a function that first rotates the image $I_j$ by $\theta$ degrees, then translates the image by $dx$ pixels in the $x$ direction, and finally translates the image by $dy$ pixels in the $y$ direction. 
%
$\Delta$ is the maximum number of pixels by which we are allowed to translate the image (we take $\Delta=20$, which corresponds to a 20\% shift in the image). 
%
The solution to \eqref{eq:opt_pairwise} is not easily computed, as the objective function will most likely be nonconvex.
%
Therefore, instead of using an optimization procedure to compute the solution to \eqref{eq:opt_pairwise}, we choose to discretize the search space and exhaustively search to estimate the solution to \eqref{eq:opt_pairwise}.
%
We discritize the search space of rotations into 20 possible rotations ($0, \pi/10, \pi/5, \dots, 9 \pi/5, 19\pi/10$), and 11 possible translations in both the $x$ and $y$ directions ($-20, -16, -12, \dots, 12, 16, 20$ pixels). 
%
We then check all possible combinations for the best rotation and translations that align $I_i$ to $I_j$. 
%
Although this can be somewhat time intensive, it is not prohibitive for the size of data sets which we are considering, and can be trivially parallelized for larger data sets if necessary.
%
The solution to this search will not be the exact solution to \eqref{eq:opt_pairwise}, but it will (most likely) be a close approximation.
%
Since our techniques are robust to noise, close approximations will be sufficient to obtain accurate results.

@@

We now compute the three--dimensional rotation, $R_{ij} \in SO(3)$, from the two--dimensional translation and rotation parameters $\theta_{ij}$, $dx_{ij}$, and $dy_{ij}$.
%
In the space of three--dimensional rotations, the Euler angles $\alpha_{ij}$, $\beta_{ij}$, and $\gamma_{ij}$ correspond to
\begin{equation} \label{eq:angle_relations}
\begin{aligned}
	\alpha_{ij} &= \theta_{ij} \\
	\beta_{ij} &= \frac{dx_{ij}}{n} \times \eta_{proj} \\
	\gamma_{ij} &= \frac{dy_{ij}}{n} \times \eta_{proj} \\
\end{aligned}
\end{equation}
where $\eta_{proj}$ is the angular portion of the sphere onto which we choose to project (we take $\eta_{proj} =  \pi/8$).
%
From here, we can write rotations around the three principal axes in terms of the three Euler angles
\begin{equation}
\begin{aligned}
	R^x_{ij} &= \begin{bmatrix}
	1 & 0 & 0 \\
    0 & \cos(\alpha) & -\sin(\alpha) \\
    0 & \sin(\alpha) & \cos(\alpha)
	\end{bmatrix} \\
	R^y_{ij} &= \begin{bmatrix}
	\cos(\beta) & 0 & \sin(\beta) \\
    0 & 1 & 0 \\
    -\sin(\beta) & 0 & \cos(\beta)
    \end{bmatrix} \\
	R^z_{ij} &= \begin{bmatrix} 
	\cos(\gamma) & -\sin(\gamma) & 0 \\
    \sin(\gamma) & \cos(\gamma) & 0 \\
    0 & 0 & 1 
    \end{bmatrix}
\end{aligned}
\end{equation}
where $R^x_{ij}$, $R^y_{ij}$, and $R^z_{ij}$ correspond to rotations around the $x$, $y$, and $z$ axes, respectively.
%
We then write the total rotation $R_{ij} \in SO(3)$ as 
\begin{equation} \label{eq:total_rot}
	R_{ij}	 = R^z_{ij} \times R^y_{ij} \times R^x_{ij}
\end{equation}
%
\eqref{eq:total_rot} corresponds to first rotating the image by $\theta$, then translating the image by $dx$ pixels in the $x$ direction, and finally translating the image by $dy$ pixels in the $y$ direction (we note that, because the rotation matrices operate from the left, the rightmost rotation matrix in the product in \eqref{eq:total_rot} corresponds to the first operation we perform on the image).

@@

Because we relaxed the problem to allow our solutions to lie in $\mathbb{R}^{3 \times 3}$, we must project our approximate solution back to $SO(3)$.
%
The optimal rotations are given by $R_i = U_i V_i^T$, where $U_i$ and $V_i$ are the left and right singular vectors of $\hat{R}_i$, respectively \cite{...}. 
%
Because the eigenvectors of $H$ are determined up to a sign, we adjust the signs of the eigenvectors so that $det(R_i) = +1$ to ensure proper rotations.
%

@@

\section{Example: {\em Drosophila}}

\subsection{dpERK}

\begin{figure*}[h]
\begin{subfigure}[c]{0.45\textwidth}
\includegraphics[width=\textwidth]{dpERK_array}
\caption{}
\end{subfigure}
\begin{subfigure}[c]{0.5\textwidth}
\includegraphics[width=\textwidth]{dpERK_rank_corr}
\caption{}
\end{subfigure}
\caption{(a) Ordering of images fluorescently stained for dpERK. (b) Correlation between ranks from ordering using VDM, and ordering using the membrane length.}
\label{fig:dpERK}
\end{figure*}

\begin{figure*}[h]
\centering
\includegraphics[width=0.5\textwidth]{bootstrap_rank_corr}
\caption{Rank correlation as a function of the number of data points.}
\label{fig:bootstrap}
\end{figure*}

\subsection{Dorsal}

\begin{figure*}[h]
\begin{subfigure}[c]{0.45\textwidth}
\includegraphics[width=\textwidth]{DI_array}
\caption{}
\end{subfigure}
\begin{subfigure}[c]{0.5\textwidth}
\includegraphics[width=\textwidth]{DI_rank_corr}
\caption{}
\end{subfigure}
\caption{(a) Ordering of images fluorescently stained for Dorsal. (b) Correlation between ranks from ordering using VDM, and ordering using the membrane length.}
\label{fig:Dl}
\end{figure*}

\section{Discussion}
% compare to TSP algorithms
Discussion of what we did and how it relates to other efforts (TSP, etc) and how it can be extended.

>>>>

@@

@@

@@

Let $I_1, \dots, I_m \in \mathbb{R}^{n \times n}$ denote the (grayscale) images that we wish to align with respect to translations and rotations.
%
The first step in angular synchronization is to compute the translations and rotations to optimally align each pair of data points. 
%
For each image pair $I_i$ and $I_j$, we would like to solve
\begin{equation}\label{eq:opt_pairwise}
(\theta_{ij}, dx_{ij}, dy_{ij}) = \argmin_{
\begin{matrix}
\theta \in [0, 2\pi) \\ 
dx \in [-\Delta, \Delta]\\ 
dy \in [-\Delta, \Delta]
\end{matrix}
} \|f(I_j, \theta, dx, dy) - I_i \|_F^2.
\end{equation}
where $\| \cdot \|_F$ denotes the Frobenius norm, and $f(I_j, \theta, dx, dy)$ is a function that first rotates the image $I_j$ by $\theta$ degrees, then translates the image by $dx$ pixels in the $x$ direction, and finally translates the image by $dy$ pixels in the $y$ direction. 
%
$\Delta$ is the maximum number of pixels by which we are allowed to translate the image.

@@

We now compute the three--dimensional rotation, $R_{ij} \in SO(3)$, from the two--dimensional translation and rotation parameters $\theta_{ij}$, $dx_{ij}$, and $dy_{ij}$.

@@

Because we relaxed the problem to allow our solutions to lie in $\mathbb{R}^{3 \times 3}$, we must project our approximate solution back to $SO(3)$.
%

@@

\section{Appendix}

\subsection{Angular Synchronization Implementation}
(we take $\Delta=20$, which corresponds to a 20\% shift in the image). 
%
The solution to \eqref{eq:opt_pairwise} is not easily computed, as the objective function will most likely be nonconvex.
%
Therefore, instead of using an optimization procedure to compute the solution to \eqref{eq:opt_pairwise}, we choose to discretize the search space and exhaustively search to estimate the solution to \eqref{eq:opt_pairwise}.
%
We discritize the search space of rotations into 20 possible rotations ($0, \pi/10, \pi/5, \dots, 9 \pi/5, 19\pi/10$), and 11 possible translations in both the $x$ and $y$ directions ($-20, -16, -12, \dots, 12, 16, 20$ pixels). 
%
We then check all possible combinations for the best rotation and translations that align $I_i$ to $I_j$. 
%
Although this can be somewhat time intensive, it is not prohibitive for the size of data sets which we are considering, and can be trivially parallelized for larger data sets if necessary.
%
The solution to this search will not be the exact solution to \eqref{eq:opt_pairwise}, but it will (most likely) be a close approximation.
%
Since our techniques are robust to noise, close approximations will be sufficient to obtain accurate results.

%
In the space of three--dimensional rotations, the Euler angles $\alpha_{ij}$, $\beta_{ij}$, and $\gamma_{ij}$ correspond to
\begin{equation} \label{eq:angle_relations}
\begin{aligned}
	\alpha_{ij} &= \theta_{ij} \\
	\beta_{ij} &= \frac{dx_{ij}}{n} \times \eta_{proj} \\
	\gamma_{ij} &= \frac{dy_{ij}}{n} \times \eta_{proj} \\
\end{aligned}
\end{equation}
where $\eta_{proj}$ is the angular portion of the sphere onto which we choose to project (we take $\eta_{proj} =  \pi/8$).
%
From here, we can write rotations around the three principal axes in terms of the three Euler angles
\begin{equation}
\begin{aligned}
	R^x_{ij} &= \begin{bmatrix}
	1 & 0 & 0 \\
    0 & \cos(\alpha) & -\sin(\alpha) \\
    0 & \sin(\alpha) & \cos(\alpha)
	\end{bmatrix} \\
	R^y_{ij} &= \begin{bmatrix}
	\cos(\beta) & 0 & \sin(\beta) \\
    0 & 1 & 0 \\
    -\sin(\beta) & 0 & \cos(\beta)
    \end{bmatrix} \\
	R^z_{ij} &= \begin{bmatrix} 
	\cos(\gamma) & -\sin(\gamma) & 0 \\
    \sin(\gamma) & \cos(\gamma) & 0 \\
    0 & 0 & 1 
    \end{bmatrix}
\end{aligned}
\end{equation}
where $R^x_{ij}$, $R^y_{ij}$, and $R^z_{ij}$ correspond to rotations around the $x$, $y$, and $z$ axes, respectively.
%
We then write the total rotation $R_{ij} \in SO(3)$ as 
\begin{equation} \label{eq:total_rot}
	R_{ij}	 = R^z_{ij} \times R^y_{ij} \times R^x_{ij}
\end{equation}
%
\eqref{eq:total_rot} corresponds to first rotating the image by $\theta$, then translating the image by $dx$ pixels in the $x$ direction, and finally translating the image by $dy$ pixels in the $y$ direction (we note that, because the rotation matrices operate from the left, the rightmost rotation matrix in the product in \eqref{eq:total_rot} corresponds to the first operation we perform on the image).

The optimal rotations are given by $R_i = U_i V_i^T$, where $U_i$ and $V_i$ are the left and right singular vectors of $\hat{R}_i$, respectively \cite{...}. 
%
Because the eigenvectors of $H$ are determined up to a sign, we adjust the signs of the eigenvectors so that $det(R_i) = +1$ to ensure proper rotations.
%
